---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

```{r, include = FALSE}
# library(ccmc)
devtools::load_all()
library(tibble)
library(ggplot2)
library(tidyr)
library(dplyr)
```

# ccmc
Cluster Correction for Multiple Comparisons (v. 0.0.0.9000)

The goal of ccmc is to correct for multiple comparisons of dependent tests using cluster-based statistics. The rationale for this approach is described [here](https://garstats.wordpress.com/2018/09/06/cluster/).

## Installation

The package can be installed using these commands:
```{r, eval = FALSE}
install.packages("devtools")
devtools::install_github("GRousselet/ccmc")
```


## Examples

### No effect: 5 independent groups

```{r}
set.seed(21)
N <- 40 # observations per group
Ng <- 5 # number of groups
x <- matrix(rnorm(N*Ng), ncol = Ng)
trimbt.ccmc(x,nullval=0,tr=0,alpha=.05,bt=FALSE,nboot=599)
```

### Cluster of length 3 with 5 independent groups

Make data
```{r}
set.seed(21)
N <- 40 # observations per group
Ng <- 5 # number of groups
x <- matrix(rnorm(N*Ng), ncol = Ng)
x[,3:5] <- x[,3:5] + 1
```

Illustrate data
```{r}
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

# data frame of independent observations
df <- as.tibble(x)
colnames(df) <- c("c1", "c2", "c3", "c4", "c5")
df <- gather(df, condition, value)
df$condition <- factor(df$condition)

# data frame of means
tmp <- as_tibble(x)
colnames(tmp) <- c("c1", "c2", "c3", "c4", "c5")
df.mean <- tidyr::gather(dplyr::summarise_all(tmp, funs(mean)),condition, value)
df.mean$condition <- as.factor(df.mean$condition)

ggplot(df, aes(x=condition, y=value, colour=condition)) + theme_grey() + 
  geom_jitter() +
  scale_colour_manual(values=cbPalette) +
  geom_line(data = df.mean, aes(group = 1), colour = "black", size = 1) +
  geom_point(data = df.mean, colour = "black", size = 3) +
  geom_abline(slope = 0, intercept = 0) + 
  theme(axis.title = element_text(size = 18),
        axis.text = element_text(size = 14, colour="black"),
        legend.position="none")
```

Mass cluster test
```{r}
trimbt.ccmc(x,nullval=0,tr=0,alpha=.05,bt=FALSE,nboot=599)
```


Use bootstrap-t thresholds
```{r, eval = FALSE}
trimbt.ccmc(x,nullval=0,tr=0,alpha=.05,bt=TRUE,nboot=599)
```

Get cluster statistics for cluster 1:
```{r, eval = FALSE}
out <- trimbt.ccmc(x,nullval=0,tr=0,alpha=.05,bt=TRUE,nboot=599)
c2sum <- sum(out$tval[out$cluster.map==1]^2)
```

### Cluster of length 2: 5 dependent groups

Make data
```{r}
require(MASS)
set.seed(21)
Ng <- 5 # n groups
Np <- 30 # n participants per group
rho = 0.75 # correlation between repeated measures
ES <- c(0, 1, 1, 0, 0) # true effects
sigma <- 1 # population standard deviation

# variance-covariance matrix
Sigma <- diag(nrow = Ng)
Sigma[Sigma == 0] <- rho

x <- mvrnorm(Np, ES, Sigma)
```

Illustrate data
```{r}

df <- as.tibble(x)
colnames(df) <- c("c1", "c2", "c3", "c4", "c5")
df <- gather(df, condition, value)
df$participant <- factor(rep(seq(1, Np), Ng))
df$condition <- factor(df$condition)

# data frame of means
tmp <- as_tibble(x)
colnames(tmp) <- c("c1", "c2", "c3", "c4", "c5")
# apply(data, 2, mean) # Mean per group
df.mean <- tidyr::gather(dplyr::summarise_all(tmp, funs(mean)),condition, value)
df.mean$condition <- as.factor(df.mean$condition)

ggplot(df, aes(x=condition, y=value)) +
  geom_line(aes(group=participant, colour=participant)) +
  geom_line(data = df.mean, aes(group = 1), size = 1) +
  geom_point(data = df.mean, aes(group = 1), size = 3) +
  geom_abline(slope = 0, intercept = 0) + 
  theme(axis.title = element_text(size = 18),
        axis.text = element_text(size = 14, colour="black"))
```

Mass cluster test
```{r}
trimbt.ccmc(x,nullval=0,tr=0,alpha=.05,bt=FALSE,nboot=599)
```


## Threshold Free Cluster Enhancement

TFCE demo
```{r}
x.seq <- seq(1,29)
data <- c(0,0,0,0,seq(1,11),seq(10,1),0,0,0,0)
data.tfce <- tfce(data)

df <- tibble(x = rep(x.seq,2),
             y = c(data/max(data), data.tfce/max(data.tfce)),
             data = factor(c(rep("Original",29),rep("TFCE",29)))
             )

ggplot(df, aes(x=x, y=y, colour = data)) + 
  geom_point(size = 2) +
  geom_line(size = 0.5) +
scale_colour_manual(values=c("black","orange"))
```

TFCE bootstrap data
```{r}
nboot <- 1000
Ng <- 5
boot.data <- matrix(rnorm(Ng*nboot), ncol = nboot)
boot.tfce <- tfce(abs(boot.data))
```

Correct for multiple comparisons using TFCE
```{r}
set.seed(21)
N <- 40 # observations per group
Ng <- 5 # number of groups
x <- matrix(rnorm(N*Ng), ncol = Ng)
x[,3:5] <- x[,3:5] + 1
trimbt.tfce(x,nullval=0,tr=0,alpha=.05,nboot=599)
```

